Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-03-17 paper/manuscript.Rmd
Update manuscript.Rmd

For several decades after their introduction to Fogo Island, the caribou
population did not exceed ~100 individuals [@Bergerud1989, Newfoundland and
Labrador Wildlife Division, unpublished data]. During the 1990s, population
density reached 300 animals and anecdotal evidence suggests that
caribou began to occupy other islands in the archipelago. We surmise that as
competition among conspecifics increased along with population size,
density-dependent habitat selection resulted in expansion of the population to
nearby islands. Following Ideal Free Distribution theory, caribou should swim to
new islands when the average fitness of individuals on the starting island
exceeds the density-fitness equilibrium [@Morris1987]. While this is an ultimate
explanation to a series of proximate observations, fine-scale density-dependent
habitat selection is theoretically possible [@Webber2018]. Given the presumed
low density of coyotes on Fogo Island and the minimal risk they pose to all but
young calves, who are least likely to engage in oceanic swimming, predation risk
does not appear to be a major driver of the behaviour in this system.
<!-- QW: changed it a bit - how is this? JGH: simplified these two para's a bit,
took out most of the predator stuff. I'm a bit confused by the sentence above
citing your 2018 paper, I'm not really sure what it's saying & how it relates to
the rest of the section. there were comments to that effect as well, maybe
expand the end a bit to make it clearer how that relates? -->

>>>>

For several decades after their introduction to Fogo Island, the caribou
population did not exceed ~100 individuals [@Bergerud1989, Newfoundland and
Labrador Wildlife Division, unpublished data]. During the 1990s, population
density reached 300 animals and it is unknown when caribou began to occupy other
islands, anecdotal evidence suggests it was during the 1990s. We surmise that as
competition among conspecifics increased along with population size,
density-dependent habitat selection resulted in expansion of the population to
nearby islands. Following Ideal Free Distribution theory, caribou should swim to
new islands when the average fitness of individuals on the starting island
exceeds the density-fitness equilibrium [@Morris1987]. While this is an ultimate
explanation to a series of proximate observations, fine-scale density-dependent
habitat selection is theoretically possible [@Webber2018]. Given the presumed
low density of coyotes on Fogo Island and the minimal risk they pose, predation risk
does not appear to be a major driver of swimming.

<!-- QW: changed it a bit - how is this? JGH: simplified these two para's a bit,
took out most of the predator stuff. I'm a bit confused by the sentence above
citing your 2018 paper, I'm not really sure what it's saying & how it relates to
the rest of the section. there were comments to that effect as well, maybe
expand the end a bit to make it clearer how that relates? -->
